---
title: "Darwin Core Mapping"
author: "Dimitri Brosens"
output: html_document
date: "`r Sys.Date()`"

---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
##library(rgbif)        # To use GBIF services
##library(uuid)         # To generate UUID's
library(sp)             # coordinate transformation
library(sf)             #coordinate transformation
```

# Read source data



```{r}
input_data <- read_delim(here::here("datasets", "atlas_reptilia","data","raw", "20230321_reptilia.csv"), ",", locale = locale(encoding = "windows-1252") )
```


Preview data:

```{r}
input_data %>% head()
```



# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% 
  remove_empty("rows") %<>%
  remove_empty("cols") %<>%
  clean_names()
```


## remove obsolete columns

```{r}
input_data <- input_data %>% select(-c(
                                email
                              , reproductivecondition
                              , dateentered
                              , firstname
                              , lastname
                              , dateentered,
                              , municipalityid
                              , municipality_id
                              , postalcode
                              , isinbcr
                              , parent_id
                              , accepted_id
                              , gbif_tree_top))
  
```


### fix unique ID issue

```{r}
input_data <- mutate_all(input_data, as.character())
```
# Georeference source data

## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates

We want to keep the original verbatim Coordinates
The original X,Y will be used for creating decimalLatitude and decimalLongitude

```{r}
input_data %<>% mutate(Longitude = lambertlongitude) %>%
                   mutate(Latitude = lambertlatitude)
```
### remove columns with NA in lat or long

```{r}
input_data %<>% drop_na(Longitude)
```



### Create a spacial dataframe

Create spatial dataframe and define the CRS (31370 is Belgian Lambert 72)(crs = 28992 is the Dutch system used here)
We indicate the x,y columns as coordinates and define the coordinate system

```{r}
input_data = st_as_sf(input_data, coords = c("Longitude","Latitude"), crs = 31370 , agr = "constant" )
input_data %>% head(n = 5)
```

## Transform the data in WGS84

Now we transform the Lambert 72 in WGS84

```{r}
input_data <- st_transform(input_data, crs = 4326)
input_data %>% head(n = 5)
```

### Create a dataframe with column decimalLatitude & decimalLongitude

In order to add the decimal coordinates into the original df we create an x_y dataframe

```{r}
x_y <- as.data.frame(st_coordinates(input_data))
x_y %>% head(n = 5)
write_csv(x_y, here::here("datasets", "atlas_reptilia", "data", "interim/x_y.csv"), na = "")
```

# Create extra dataframe to work with 

```{r}
input_interim_df <- input_data
```

# Remove geometry from dataframe 'input_interim_df'

We remove all geometry from the spacial dataframe, to create a normal df

```{r}
##input_interim_df <- st_geometry(NULL)
st_geometry(input_interim_df) <- NULL
input_interim_df %>% head(n = 5)
```

### Bind columns x_y and input_interim_df

Join both df's to have the decimalLat & Lon

```{r}
input_interim_df = bind_cols(input_interim_df,x_y, .id = NULL)
input_interim_df %>% head(n = 5)
```

### Integrate x-y coordinates in original dataframe

We can safely substitute input_interim 

```{r}
input_data <- input_interim_df
```



# Darwin Core mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
occurrence <- input_data
```

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO")
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/daf62d")
```

### institutionCode


### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Invasive species - American bullfrog (Lithobates catesbeianus) in Flanders, Belgium (post 2018)")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### occurrenceID


```{r}
occurrence %<>% mutate(dwc_occurrenceID = occurrence_id)
```




### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = recorded_by)
```

### individualCount

```{r}
#mutate(occurrence$mm)as.numeric()
  
occurrence %<>% mutate(dwc_individualCount = '')

```




### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = 'present')
```

```{r}
occurrence %<>% mutate(dwc_occurrenceRemarks = 'bycatch')
```


### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = str_replace_all(event_id, " ",""))
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = datum)
```




### habitat

```{r}
occurrence %<>% mutate(dwc_habitat = recode(type
                                            , 'Stilstaand' = 'freshwater; stagnant'))
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = recode(actie
                                                     ,"Afvangst" = "catch, fike"
                                                     
                                                     )) 
```
### sampleSizeValue

```{r}
occurrence %<>% mutate(dwc_sampleSizeValue = no_fuiken) 
```
### sampleSizeUnit

```{r}
occurrence %<>% mutate(dwc_sampleSizeUnit = recode(fuiktype
                                                   ,"dubbele fuik" = "paired fyke"
                                                   ,"salamanderfuik drijvend NA" = "amphibian trap floating"
                                                   ,"salamanderfuik bodem" = "amphibian trap bottom"
                                                   ,"enkele fuik" = "single fyke"
                                                   ,"vrijwilligers Natuurpunt" = "unspecified fyke net"
                                                   , .default = "fyke"
                                                     )) 
```

### sex

```{r}
occurrence %<>% mutate(dwc_sex = 'unknown')
```

### lifeStage

```{r}
occurrence %<>% mutate(dwc_lifeStage = 'unknown')

#ask Sander for info

```


### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = location_id)
```


### locality

```{r}

occurrence %<>% mutate(dwc_locality = paste(locatie))


```


### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe")
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = 'BE')
```



### locationRemarks



### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.5f",
  round(lat, digits = 5)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.5f",
  round(long, digits = 5)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84") 
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```
### Identification


```{r}
occurrence %<>% mutate(dwc_identifiedBy = identified_by)
```

```{r}
#occurrence %<>% mutate(dwc_identifiedByID = "https://orcid.org/0000-0001-5198-5928")
```


```{r}
#occurrence %<>% mutate(dwc_dateIdentified = identified_date)
```




### scientificName

### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = recode(scientificname
                                                     ,"Distycus spec." = "Distycus"
                                                      ))

```


```{r}
occurrence %<>% filter(!is.na(dwc_scientificName))

#df %>% filter(!is.na(col1))
```


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### vernacularName


### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = case_when(dwc_scientificName == 'Distycus' ~ 'genus'
                                                 , TRUE ~ "species"))
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

## Post-processing

```{r}
occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

```{r}
occurrence$eventDate <- str_replace_all(occurrence$eventDate, "NA", "")
```


```{r}
#occurrence <-  mutate_at(c(occurrence$eventDate), ~replace_na(.,""))

```



```{r}
#occurrence %<>% distinct(occurrenceID2, .keep_all = TRUE)
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("darwincore", "processed", "occurrenceSKBC.csv"), na = "")
```

# Checks

```{r}
test <- length(unique(occurrence$eventID))
test2 <- length(unique(occurrence$occurrenceID))
n_occur <- data.frame(table(occurrence$occurrenceID))
traps <- distinct(input_data, datasetname, .keep_all = FALSE)
traps2 <- distinct(occurrence, sampleSizeUnit, .keep_all = FALSE)

```


